---
title: 'Quiz: Diamonds'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

a. Load the 'diamonds' data set in R Studio. How many observations are in the data set?

```{r}
library(ggplot2)
data("diamonds")
```

```{r}
summary(diamonds)
```

```{r}
dim(diamonds)
```

53940 observations

b. How many variables are in the data set?

-> 10

c. How many ordered factors are in the data set?

```{r}
str(diamonds)
```

-> 3

d. What letter represents the best color for a diamond?

```{r}
levels(diamonds$color)
```

J

```{r}
?diamonds
```

## Price Histogram

Create a histogram of the price of all the diamonds in the diamond data set.

```{r}
qplot(data=diamonds, x=price, binwidth=500,
      color = I('black'), fill = I('#099DD9'))
```

## Price Histogram Summary

Describe the shape and center of the price distribution. Include summary statistics like the mean and median.

```{r}
summary(diamonds)
```

Median = 3933
Median = 2401

## Diamond Counts

a. How many diamonds cost less than $500?
```{r}
subset(diamonds, price < 500)
```

b. How many diamonds cost less than $250?
```{r}
subset(diamonds, price < 250)
```

c. How many diamonds cost $15,000 or more?
```{r}
subset(diamonds, price >= 15000)
```

Explore the largest peak in the price histogram you created earlier. Try limiting the x-axis, altering the bin width, and setting different breaks on the x-axis. There won't be a solution video for this question so go to the discussions to share your thoughts and discover what other people find. You can save images by using the ggsave() command. ggsave() will save the last plot created. For example:

qplot(x = price, data = diamonds)
ggsave('priceHistogram.png')

ggsave currently recognises the extensions eps/ps, tex (pictex), pdf, jpeg, tiff, png, bmp, svg and wmf (windows only).

```{r}
qplot(data=diamonds, x=(price/1000), binwidth=0.25,
      color = I('black'), fill = I('#099DD9')) +
  scale_x_continuous(limits = c(0, 15), breaks = seq(0, 15, 2))
```

```{r}
qplot(data=diamonds, x=(price/1000), binwidth=0.2,
      color = I('black'), fill = I('#099DD9')) +
  scale_x_continuous(limits = c(0, 5), breaks = seq(0, 5, 0.5))
```

```{r}
qplot(data=diamonds, x=log10(price), binwidth = 0.02,
      color = I('black'), fill = I('#099DD9')) +
  scale_x_continuous(limits = c(2.3, 4.5), breaks = seq(0, 5 , 0.5))
```

```{r}
ggplot(aes(x = price), data = diamonds) +
  geom_histogram(color = 'black', fill = '#F79420') +
  scale_x_log10()
```

Observations:

There are no diamonds that cost $1500.

For diamonds that cost less than $2,000, the most common price of a diamond is around $700 with the mode being $605 (binwidth = 1). The second insight comes from Paula F. Thanks Paula!
Thank you for your submission!

```{r}
qplot(x=price, data=diamonds, binwidth=50,
      color = I('black'), fill = I('#099009')) +
  scale_x_continuous(limits=c(0,3000), breaks = seq(0,3000,500))
                                   
```

## Price by cut Histograms

Break out the histogram of diamond prices by cut.

You should have five histograms in separate panels on you resulting plot.

```{r}
library(gridExtra)

p1 <- ggplot(aes(x=price), data = subset(diamonds, diamonds$cut == 'Fair')) +
  geom_histogram()
p2 <- ggplot(aes(x=price), data = subset(diamonds, diamonds$cut == 'Good')) +
  geom_histogram()
p3 <- ggplot(aes(x=price), data = subset(diamonds, diamonds$cut == 'Very Good')) +
  geom_histogram()
p4 <- ggplot(aes(x=price), data = subset(diamonds, diamonds$cut == 'Premium')) +
  geom_histogram()
p5 <- ggplot(aes(x=price), data = subset(diamonds, diamonds$cut == 'Ideal')) +
  geom_histogram()

grid.arrange(p1,p2,p3,p4,p5, ncol=1)
```

Faceting:

```{r}
qplot(x=price, data=diamonds) + 
  facet_wrap(~cut, ncol = 1)
```

Frequency polygons:

```{r}
ggplot(aes(x=price), data=diamonds) +
  geom_freqpoly(aes(color=cut))
```

## Price by Cut

1. Which cut has as the highest priced diamond?

```{r}
by(diamonds$price, diamonds$cut, summary)
```

Try again:

Try using max as the function inside of the by command. The summary function does some rounding.

You could also use digits = max(getOption('digits')) as a fourth argument in the by command. Any arguments after the function in the by command get passed to that function.

```{r}
by(diamonds$price, diamonds$cut, max)
```

- premium only

```{r}
by(diamonds$price, diamonds$cut, median)
```

## Scales and Multiple Histograms

In the last two exercises, we looked at the distribution for diamonds by cut.

Run the code below in R Studio to generate the histogram as a reminder.

```{r}
qplot(x = price, data = diamonds) + facet_wrap(~cut)
```

In the last exercise, we looked at the summary statistics for diamond price by cut. If we look at the output table, the median and quartiles are resonably close to each other.

```{r}
by(diamonds$price, diamonds$cut, summary)
```

This means that the distributions should be somewhat similar, but the histograms we created don't show that.

The 'Fair' and 'Good' diamonds appear o have different distributions compared to the better cut diamonds. They seem somewhat uniform on the left with long tails on the right.

Let's look into this more.

Look up the documentation for facet_wrap in R studio. Then, scroll back up and add a parameter to facet_wrap so that the y-axis in the histograms is not fixed. You want the y-acis to be different for each histogram.

```{r}
?facet_wrap
```

See also:

http://www.cookbook-r.com/Graphs/Facets_(ggplot2)/

```{r}
qplot(x = price, data = diamonds) + facet_wrap(~cut, scale='free_y')

```

Are your y-axis scales different on some of the panels?

It's a good idea to consider the scales when you create plots.

Don't worry if the mean and median prices of diamonds seem unintuitive right now. In later lessons, we'll learn the tools to understand what's causing these patterns.

## Price per Carat by Cut

Create a histogram of price per carat and facet it by cut. You can make adjustments to the code from the previous exercise to get started. 

Adjust the bin width and transform the scale of the x-axis using log10.

```{r}
qplot(x = price/carat, data = diamonds, color = I('black'), fill = I('blue'), 
      binwidth = 0.05) + facet_wrap(~cut, scale='free_y') +
  scale_x_log10()
```

Did you find normal-ish distributions and a slightly bimodal distribution for Very Good cuts? (If the Very Good facet looks smooth, try decreasing your bin width! If the data looks skewed, did you remember to perform a log10 transform?)

## Price Box Plots

Investigate the price of diamonds using box plots, numerical summaries, and one of the following categorical variables: cut, clarity, or color.

There won't be a solution video for this exercise so go to the discussion thread for either BOXPLOTS BY CLARITY, BOXPLOT BY COLOR, or BOXPLOTS BY CUT to share your thoughts and to see what other people found.

You can save images by using the ggsave() command. 

### Boxplots by Cut

```{r}
ggplot(data = diamonds, aes(x = cut, y = price)) +
  geom_boxplot() 
#coord_cartesian(ylim = c(0, 7000))
```

### Boxplots by Clarity
```{r}
ggplot(data = diamonds, aes(x = clarity, y = price)) +
  geom_boxplot()
#coord_cartesian(ylim=c(0, 7000))
```

### Boxplots by Color
```{r}
ggplot(data = diamonds, aes(x = color, y = price)) +
  geom_boxplot() 
#coord_cartesian(ylim = c(0, 8000))
```

Oddities observed:

More expensive ones are the ones that aren't ideal in quality. For example, color J diamonds are more expensive than colorless diamonds.

## Interquartile range - IQR

a. What is the price range for the middle 50% of diamonds with color D?

```{r}
by(diamonds$price, diamonds$color, summary) 
```

first quartile (25%) $911
third quartile (75%) $4214

b. What is the price range for the middle 50% of diamonds with color J?

first quartile (25%) $1860
third quartile (75%) $7695

c. What is the IQR for diamonds with the best color?

```{r}
IQR(subset(diamonds, color == 'D')$price)
```

d. What is the IQR for diamonds with the worst color?

```{r}
IQR(subset(diamonds, color == 'J')$price)
```

## Price per Carat Box Plots by Color

Investigate the price per carat of diamonds across the different colors of diamonds using boxplots. 

```{r}
qplot(y=price/carat, x=clarity, data=subset(diamonds, !is.na(price)), fill=color, geom="boxplot")
```

https://discussions.udacity.com/t/price-boxplots-by-color/119304/26?u=alona_11249598748233
```{r}
qplot(x = color, y=price / carat , data = diamonds, geom='boxplot', color = color, fill= color, alpha=.5) + facet_wrap(~cut)
```
(https://discussions.udacity.com/t/problem-set-3-price-per-carat-box-plots-by-color/28325/11?u=alona_11249598748233)

My solution:
```{r}
ggplot(data = diamonds, aes(x=color, y = price / carat, fill = color)) +
  geom_boxplot() +
  facet_wrap(~cut)
```



